---
title: "Data Analytics Project"
author: "Saikat Roy, Vijayesh Kumar Das, Pooja Bhatia"
date: "28 July 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(pacman)
p_load(tidyverse, data.table, stringr)

```

Changing directory and loading the 'Caracteristics' CSV file

```{r}
setwd("C:\\Users\\Saikat-PC\\Desktop\\accidents-in-france-from-2005-to-2016")
data <- read.csv("caracteristics.csv", header = TRUE)
colnames(data)

#PLEASE NOTE THAT "users" and "places" dataset have been loaded later but not displayed
```

PLEASE NOTE: "users" and "places" dataset have been loaded later but not displayed

```{r}
#Structure of data frame
colnames(data)
str(data)
```

```{r}
# Short glimpse of data set
glimpse(data)
```

# Yearly Variations of Accidents over Months (1-12)
```{r}
sample = sample_n(data,100000)

data %>% mutate(factorYear=as.factor(2000+an)) %>% 
  group_by(factorYear, mois) %>% summarise(Count=n()) %>% 
  ggplot(aes(x = mois, color=factorYear, y=Count))+ 
  ggtitle("Yearly Variations of Accidents over Months")+
  geom_line() + xlab("Months from 1-12") 
```

# Monthly Variations of Accidents over Years (2005-2016) 
```{r}
data %>% mutate(factorMonth=as.factor(mois)) %>% 
  group_by(factorMonth, an) %>% summarise(Count=n()) %>% 
  ggplot(aes(x = an+2000, color=factorMonth, y=Count))+ geom_line() + 
  xlab("Years from 2005-2016")+
  ggtitle("Monthly Variations of Accidents over Years")
```

# Hourly Distribution of Accidents
```{r}
data %>% mutate(years=as.factor(2000+an)) %>%
  #group_by(factorYear, hrmn) %>% summarise(Count=n()) %>% 
  ggplot(aes(x = hrmn, color=years))+ 
  geom_histogram(bins=48)+ xlab("Hours in 24 hours format") + 
  ggtitle("Hourly Distribution of Accidents")+
  scale_x_continuous(breaks = seq(0, 2400, by = 300))+
  ggtitle("Hourly Distribution of Accidents")

```

#Distribution of accidents by lighting conditions in which the accident occurred
```{r}
data %>% mutate(years=as.factor(2000+an)) %>%
ggplot(aes(x = lum, color=years)) + ggtitle("Distribution of accidents by lighting conditions in which the accident occurred") +xlab("Lumination") + ylab("count")+geom_bar()+
scale_x_continuous(breaks = 1:5,labels=c("Full day", "Twilight or dawn", "Night without public lighting","Night with public lighting not lit","Night with public lighting on"))+
theme(axis.text.x = element_text(angle = 15))
print(plot)

```
# Distribution of accidents by Category of road
```{r}
data1<-read.csv("places.csv", header = TRUE)
data1$year<-data[match(data$Num_Acc , data1$Num_Acc), "an"]
data1$catr<-as.factor(data1$catr)
data1%>% filter(data1$catr!=0) %>% mutate(years=as.factor(2000+year)) %>%
ggplot(aes(x =catr, color=years)) + ggtitle("Distribution of accidents by Category of road") +xlab("Category of road") + ylab("count")+geom_bar()+
scale_x_discrete(breaks = c("1","2","3","4","5","6","9"),labels=c("Highway","NationalRoad","Departmental Road","Communal Way","Off public network","Off public network","other"))+
theme(axis.text.x = element_text(angle = 15))
```


# Distribution of accidents by Atmospheric conditions

```{r}
data %>% mutate(years=as.factor(2000+an)) %>%
ggplot(aes(x = atm, color=years)) + ggtitle("Distribution of accidents by Atmospheric conditions") +xlab("Atmospheric conditions") + ylab("count")+geom_bar()+
scale_x_continuous(breaks = 1:9,labels=c("Normal","Light rain","Heavy rain","Snow - hail","Fog - smoke","Strong wind - storm","Dazzling weather","Cloudy weather","Other"))+
theme(axis.text.x = element_text(angle = 15))
```
# Distribution of accidents by surface condition

```{r}
data1$surf<-as.factor(data1$surf)
data1 %>% filter(data1$surf!=0) %>% mutate(years=as.factor(2000+year)) %>%
ggplot(aes(x =surf, color=years)) + ggtitle("Distribution of accidents by surface condition") +
  xlab("surface condition") + ylab("count")+geom_bar()+
scale_x_discrete(breaks = c("1","2","3","4","5","6","7","8"),labels=c("normal"," wet","puddles","flooded","snow","mud","icy","fat-oil"))+
theme(axis.text.x = element_text(angle = 15))
```
#Accidents according to the age
```{r}
users<-read.csv("users.csv", header = TRUE)

users$accidentYear <- data[match(users$Num_Acc , data$Num_Acc), "an"] 
users$age <- (as.integer(users$accidentYear)+2000) - users$an_nais

ggplot(users, aes(x=age, fill = NULL))+
  geom_area(stat = "bin")+
  xlab("age")+
  ylab("Total Count")+ggtitle("Distribution of Accidents over Ages")
```

